Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-06-29 doc/RMD/mlg-mcg.Rmd
add mlg-table plot (with insets!

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  mutate(MLG = as.character(MLG)) %>%
  mutate(MCG = as.character(MCG)) %>%
  as_data_frame() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  arrange(MCG) %>%
  group_by(MCG) %>%
  mutate(MLGs = sum(Freq > 0), 
         Samples = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data = list(data_frame(MLG = MLG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(MLGs)) %>%
  ungroup() %>%
  distinct(MCG, MLGs, Samples, Evenness, data)
mlgs <- mcgmlg %>%
  arrange(MLG) %>%
  group_by(MLG) %>%
  mutate(MCGs     = sum(Freq > 0), 
         Samples  = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data     = list(data_frame(MCG = MCG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(Samples), desc(MCGs)) %>%
  ungroup() %>%
  distinct(MLG, MCGs, Samples, Evenness, data)
mcgs
mlgs

@@

```{r barplots, fig.width = 10, fig.height = 4}
mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

@@

```{r}

>>>>

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  mutate(MLG = as.character(MLG)) %>%
  mutate(MCG = as.character(MCG)) %>%
  as_tibble() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  arrange(MCG) %>%
  group_by(MCG) %>%
  mutate(MLGs = sum(Freq > 0), 
         Samples = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data = list(data_frame(MLG = MLG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(MLGs)) %>%
  ungroup() %>%
  distinct(MCG, MLGs, Samples, Evenness, data)
mlgs <- mcgmlg %>%
  arrange(MLG) %>%
  group_by(MLG) %>%
  mutate(MCGs     = sum(Freq > 0), 
         Samples  = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data     = list(data_frame(MCG = MCG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(Samples), desc(MCGs)) %>%
  ungroup() %>%
  distinct(MLG, MCGs, Samples, Evenness, data)
mcgs
mlgs

@@

```{r barplots, fig.width = 5, fig.height = 15}
mcg_graph <- mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), count) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  coord_flip() +
  theme(panel.grid.major.y = element_blank()) +
  theme(legend.position = c(.745, .85)) +
  # theme(legend.box = "horizontal") +
  theme(aspect.ratio = 3) +
  labs(list(
    alpha = "MLG Evenness"
  ))

uneven <- mcgs %>% 
  filter(MCG == 5) %>% 
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
  geom_col(color = "grey62", fill = "grey72", width = 0.5) + 
  theme_void() +
  ggtitle("Evenness = 0.48") +
  theme(aspect.ratio = 1) +
  theme(title = element_text(hjust = 0.5, 
                             size = rel(1.25),
                             family = "Helvetica",
                             color = "grey30"))

even <- mcgs %>% 
  filter(MCG == 53) %>% 
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
  geom_col(color = "grey20", fill = "grey20", width = 0.75) + 
  theme_void() +
  ggtitle("Evenness = 1") +
  theme(aspect.ratio = 1) +
  theme(title = element_text(hjust = 0.5, 
                             size = rel(1.25),
                             family = "Helvetica",
                             color = "grey30"))

vp1 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.67, just = "right")
vp2 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.555, just = "right")

pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 12)
print(mcg_graph)
print(uneven, vp = vp1)
print(even, vp = vp2)
dev.off()

print(mcg_graph)
print(uneven, vp = vp1)
print(even, vp = vp2)
```

```{r mlg_barplots, fig.width = 4, fig.height = 10}
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), count) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  coord_flip()
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

@@

```{r, eval = FALSE}
